### 2014-01-30 docs/paper/paper.tex
constraints on models

Each encoder unit is attached to a circular buffer of size 8192
containing a list of instructions for calculating one encoded value. 
This size was chosen as a compromise between memory utilization and encoder flexibility.
Each instruction corresponds to delaying for one or more clock cycles
(in order to coordinate concurrent access to the interconnect),
followed by reading one decoded value a certain address.
The decoded value data is made available to the encoder unit three clock cycles later
(due to interconnect delays), at which time it is multiplied by an encoder weight
and added to the accumulated encoded value, initially 0.
A certain bit in the instruction field acts as a flag to indicate that the encoded value is ready to be passed on to the population unit after this
instruction finishes executing. The encoder unit then waits for all other encoders to finish in order to synchronize
access to the shared interconnect, then resets the accumulator and begins encoding the next value.

@@

\begin{itemize}
\item encoder instruction buffer size limits number of decoded values that can be encoded
\item dimensional restrictions -- 1D, 2D population units
\item four decoded values per population -- using more will create copies of populations to accommodate extra decoders
\end{itemize}

>>>>

Each encoder unit is attached to a circular buffer of size 8192 % FIXME check the size
containing a list of instructions for calculating one encoded value. 
This size was chosen as a compromise between memory utilization and encoder flexibility.
Each instruction corresponds to delaying for one or more clock cycles
(in order to coordinate concurrent access to the interconnect),
followed by reading one decoded value a certain address.
The decoded value data is made available to the encoder unit three clock cycles later
(due to interconnect delays), at which time it is multiplied by an encoder weight
and added to the accumulated encoded value, initially 0.
A certain bit in the instruction field acts as a flag to indicate that the encoded value is ready to be passed on to the population unit after this
instruction finishes executing. The encoder unit then waits for all other encoders to finish in order to synchronize
access to the shared interconnect, then resets the accumulator and begins encoding the next value.

@@

\begin{itemize}
\item encoder instruction buffer size limits number of decoded values that can be encoded

There is an artificial limitation on the complexity of the encoding operation due to the length of the instruction buffer that is used.
With a size of 8192, % FIXME check size
because each encoder must produce 1024 encoded values per timestep, a maximum of eight instructions, and therefore eight decoded values
(not necessarily distinct)
can be used in encoding. It is theoretically possible to implement an encoding operation that requires more decoded values to be used,
but the extra instructions used for this mean that other populations sharing that encoder will have to use fewer.
It is expected that the compiler will be able to detect this scenario and reorganize populations appropriately so as to avoid overfilling the instruction buffers.
\item dimensional restrictions -- 1D, 2D population units
% FIXME citation about why 2D is ``good enough''

As only one-dimensional and two-dimensional population units have been implemented, networks that require populations to compute functions
of three or more independent variables cannot be simulated with this design.
\item four decoded values per population -- using more will create copies of populations to accommodate extra decoders

The choice of using a maximum of four sets of decoder vectors per population is arbitrary; it seems to be a good balance between
population connectivity and memory usage. However, populations that have more than four outgoing connections must be treated specially.
Assuming none of the outgoing connections can be combined together (for example, if the function they decode only differs by a constant),
in order to implement such a population on the simulator, it is necessary to duplicate it and simulate it twice,
once with the first four decoder vectors and again with the remaining ones. The consequence of this is that a population with more than
four outgoing connections will be represented as two or more populations in simulation, thereby reducing the actual number of
distinct populations that can be simulated.
\end{itemize}
